### 2020-09-18 batch-hssr.tex
Splitting into header and content files

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\usepackage[english]{babel}
\usepackage[nottoc]{tocbibind}
\usepackage{color}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{float}
\newtheorem{theorem}{Theorem}[section]
\bibliographystyle{ims}
% Custom commands / shortcuts
\providecommand{\sign}{\textrm{sign}}
\providecommand{\note}[1]{\textcolor{red}{#1}}
\providecommand{\lam}{\lambda}

\title{Adaptive Hybrid Screening for Efficient Lasso Optimization}
\author{Chuyi Wang\\Department of Statistics and Actuarial Sciences\\University of Iowa
  \and
  Patrick Breheny\\Department of Biostatistics\\University of Iowa}
\date{}

\begin{document}

\maketitle

\input{main}

\bibliography{ref}

\end{document}