### 2020-12-04 batch-hssr.tex
Some edits to Section 2

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\usepackage[english]{babel}
\usepackage[nottoc]{tocbibind}
\usepackage{color}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{float}
\newtheorem{theorem}{Theorem}[section]
\bibliographystyle{ims}
% Custom commands / shortcuts
\providecommand{\sign}{\textrm{sign}}
\providecommand{\pb}[1]{\textcolor{red}{#1}}
\providecommand{\lam}{\lambda}

\title{Adaptive Hybrid Screening for Efficient Lasso Optimization}
\author{Chuyi Wang\\Department of Statistics and Actuarial Sciences\\University of Iowa
  \and
  Patrick Breheny\\Department of Biostatistics\\University of Iowa}
\date{}

\begin{document}

\maketitle

\begin{abstract}
Lasso type models are popular in statistics and machine learning, especially for variable selection in high-dimensional data. Typical practice is to tune the size of lasso penalty along a path of values. Due to modern data collection techniques, researchers need to analyse ultra high-dimension data with millions of features, which brings a great need for an efficient lasso optimization algorithm. Feature screening techniques have proven to be powerful for this need, because they can discard features and lead to a model with much less features. In this paper, we develop an adaptive hybrid screening framework where screening is carried out adpatively along the path of tuning parameter values. It reuses previous solutions in the path and avoids unhelpful heavy computations. It is a flexible framework that can be easily extended to different types of lasso models. The Ada-SSR-EDPP for the standard lasso model and the Ada-SSR-Slores for the sparse logistic model are shown as examples. We perform simulation study and real data study in a wide range of scenarios and the adaptive hybrid methods outperform other state-of-the-art method significantly and uniformly, with the greatest speedup in the most challenging scenarios.
\end{abstract}

\input{main}

\bibliography{ref}

\end{document}